---
title: "p8105_hw2_hca2124"
author: "Helene Apollon"
date: "2026-10-08"
output: github_document
---

#Question 1

```{r}
library(tidyverse)

NYC_transit_df =
  read_csv("./NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
           na = c ("NA", ".", ""))|>
janitor::clean_names() |>
  select(line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada) |>
mutate(
  entry = 
         case_match(
           entry,
           "YES" ~ TRUE,
           "NO" ~ FALSE
         ))
  
```
This data set contains information on the entrances and exits for each station in NYC. It contains `r ncol(NYC_transit_df)` variables. From the original data, I cleaned the names, and kept only the variables of interest using the select line followed by the variable names. Then I converted the entry variable from a character (Yes vs No), to a logical variable. The final dataset has `r nrow(NYC_transit_df)` rows and `r ncol(NYC_transit_df)` columns. The data is not tidy because the Route variable is spread across 11 variable columns from route1:route11.

```{r}
NYC_transit_df |> 
  distinct(line, station_name) |> 
  nrow()
```
There are `r nrow(distinct(NYC_transit_df, line, station_name))` distinct stations.

```{r}
NYC_transit_df |>
  filter(ada) |>
  distinct(line, station_name) |>
  nrow()
```
There are `r nrow(distinct(filter(NYC_transit_df, ada), line, station_name))` ADA compliant stations.

```{r}
NYC_transit_df |>
  filter(vending == "NO") |>
  pull(entry) |>
  mean()
```
`r round(mean(pull(filter(NYC_transit_df, vending == "NO"), entry)),2)` of the station entrances/exits without vending allow entrance

```{r}
NYC_transit_reformat_df =
  NYC_transit_df |>
  mutate(route8 = as.character(route8),
         route9 = as.character(route9),
         route10 = as.character(route10),
         route11 = as.character(route11)) |>
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    values_to = "route_name",
    names_prefix = "route",
    values_drop_na = TRUE
  )

# Distinct stations that serve the A
A_train =
  NYC_transit_reformat_df |>
  filter(route_name == "A") |>
  distinct(station_name, line, .keep_all = TRUE) |>
  nrow()

# A stations that are ADA compliant
ADA_A_stations = 
  NYC_transit_reformat_df |>
  filter(route_name == "A", ada) |>
  distinct(line, station_name) |>
  nrow()
  
  

```

There are `r A_train` distinct stations that serve the A train.
Of the stations that serve the A train, `r ADA_A_stations` are ADA compliant.




