---
title: "p8105_hw2_hca2124"
author: "Helene Apollon"
date: "2026-10-08"
output: github_document
---

#Question 1

```{r}
library(tidyverse)

NYC_transit_df =
  read_csv("./NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
           na = c ("NA", ".", ""))|>
janitor::clean_names() |>
  select(line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada) |>
mutate(
  entry = 
         case_match(
           entry,
           "YES" ~ TRUE,
           "NO" ~ FALSE
         ))
  
```
This data set contains information on the entrances and exits for each station in NYC. It contains `r ncol(NYC_transit_df)` variables. From the original data, I cleaned the names, and kept only the variables of interest using the select line followed by the variable names. Then I converted the entry variable from a character (Yes vs No), to a logical variable. The final dataset has `r nrow(NYC_transit_df)` rows and `r ncol(NYC_transit_df)` columns. The data is not tidy because the Route variable is spread across 11 variable columns from route1:route11.

```{r}
NYC_transit_df |> 
  distinct(line, station_name) |> 
  nrow()
```
There are `r nrow(distinct(NYC_transit_df, line, station_name))` distinct stations.

```{r}
NYC_transit_df |>
  filter(ada) |>
  distinct(line, station_name) |>
  nrow()
```
There are `r nrow(distinct(filter(NYC_transit_df, ada), line, station_name))` ADA compliant stations.

```{r}
NYC_transit_df |>
  filter(vending == "NO") |>
  pull(entry) |>
  mean()
```
`r round(mean(pull(filter(NYC_transit_df, vending == "NO"), entry)),2)` of the station entrances/exits without vending allow entrance

```{r}
NYC_transit_reformat_df =
  NYC_transit_df |>
  mutate(route8 = as.character(route8),
         route9 = as.character(route9),
         route10 = as.character(route10),
         route11 = as.character(route11)) |>
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    values_to = "route_name",
    names_prefix = "route",
    values_drop_na = TRUE
  )

# Distinct stations that serve the A
A_train =
  NYC_transit_reformat_df |>
  filter(route_name == "A") |>
  distinct(station_name, line, .keep_all = TRUE) |>
  nrow()

# A stations that are ADA compliant
ADA_A_stations = 
  NYC_transit_reformat_df |>
  filter(route_name == "A", ada) |>
  distinct(line, station_name) |>
  nrow()
  
  

```

There are `r A_train` distinct stations that serve the A train.
Of the stations that serve the A train, `r ADA_A_stations` are ADA compliant.

#Question 2

```{r}
library(readxl)
mrtrash_df = 
  read_excel("./202610 Trash Wheel Collection Data.xlsx",
             sheet = "Mr. Trash Wheel",
             range = "A2:N742") |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(sports_balls = as.integer(round(sports_balls)),
         year = as.numeric(year),
         trash_wheel = "Mr. Trash Wheel")
```

```{r}
professortrash_df =
  read_excel("./202610 Trash Wheel Collection Data.xlsx",
             sheet = "Professor Trash Wheel",
             range = "A2:M139") |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(year = as.numeric(year),
         plastic_bottles = as.numeric(plastic_bottles),
         trash_wheel = "Professor Trash Wheel")
```

```{r}
gwynnda_df =
  read_excel("./202610 Trash Wheel Collection Data.xlsx",
             sheet = "Gwynnda the Good Wheel of the W",
             range = "A2:L394") |>
  janitor::clean_names() |>
  drop_na(dumpster) |>
  mutate(year = as.numeric(year),
         trash_wheel = "Gwynnda Trash Wheel")
```

#Combining datasets
```{r}
trashwheels_df =
  bind_rows(mrtrash_df, professortrash_df, gwynnda_df) |>
  janitor::clean_names()
```

The final trashwheels dataset has a total of `r nrow(trashwheels_df)` observations. These observations are a combination of data from Mr. Trash Wheel, Gwynnda the Good Wheel of the W, and Professor Trash Wheel. 

Professor Trash Wheel collected `r filter(trashwheels_df,trash_wheel == "Professor Trash Wheel") |> select(weight_tons) |> sum()` tons of trash.

Gwynnda collected a total number of `r filter(trashwheels_df,trash_wheel =="Gwynnda Trash Wheel", month == "June", year == 2022) |> select(cigarette_butts) |> sum() |> format (scientific=FALSE)`cigarette butts in June 2022.

#Question 3
```{r}
mci_df = read_csv("./MCI_baseline.csv",
                  skip = 1,
                  na = c("NA", ".", "")) |>
  janitor::clean_names() |>
  mutate(
    sex = case_match(sex, 1 ~"male", 0 ~ "female"),
    apoe4 = case_match(apoe4, 1 ~ "carrier", 0 ~ "non-carrier"))

```

```{r}
mci_baseline =
  mci_df |>
  filter(is.na(age_at_onset) | age_at_onset > current_age)
```

The first line of the imported file was skipped as it did not contain data, rather comments.Sex and apoe4 were recoded and converted from numerical to character values. We then filterd to keep those who did not develop MCI/ MCI came after baseline age.

There were `r nrow(mci_df)` participants recruited. Of those recruited, `r mci_baseline |> filter(!is.na(age_at_onset)) |> nrow()` developed MCI. The average baseline age is `r round(mean(pull(mci_baseline, current_age)),1)` years. `r round(100 * mean(pull(filter(mci_baseline, sex == "female"), apoe4) == "carrier"), 1)` % of women in the study are apoe4 carriers.

```{r}
mci_amyloid_df =
  read_csv("mci_amyloid.csv",
           skip = 1,
           na = c("NA", ".", "")) |>
  janitor::clean_names() |>
  rename(id = study_id)
```

The first line of the imported file was skipped as it did not contain data, rather comments. 'Study_id' was renamed to 'id' in order to match the baseline data.This dataset has `r nrow(mci_amyloid_df)` participants.

```{r}
anti_join(mci_baseline, mci_amyloid_df, by = "id") |>
  nrow()

anti_join(mci_amyloid_df, mci_baseline, by = "id") |>
  nrow()

```

There are `r anti_join(mci_baseline, mci_amyloid_df, by = "id") |> nrow()` participants in the baseline that are not in the amyloid dataset. There are `r anti_join(mci_amyloid_df, mci_baseline, by = "id") |> nrow()` participants in the amyloid that are not in the baseline dataset.

```{r}
mci_combined_df =
  inner_join(mci_baseline, mci_amyloid_df, by = "id")

write_csv(mci_combined_df, "mci_combined_df.csv")
```

There are `r nrow(mci_combined_df)` subjects in the combined dataset.There is a combined total of `r ncol(mci_combined_df)` variables.




